Redshift Evolution of the Underlying Type Ia Supernova Stretch Distribution

### 2021-02-08 Nicolas.tex
Two columns for us 8-)
Rid of textbf, 5sigma in abstract, 1 column for referee

\abstract{The detailed nature of type Ia supernovae (SNe~Ia) remains uncertain,
    and as survey statistics increase, the question of astrophysical
    systematic uncertainties rises, notably that of SN~Ia population
    evolution. In this paper, we study the dependence with redshift of the SN~Ia
    \texttt{SALT2.4} lightcurve stretch, a purely intrinsic SN property, to
    probe its potential redshift drift. The SN stretch has been shown to
    strongly correlate with the SN environment, notably with stellar age
    tracers. We model the underlying stretch distribution as a function of
    redshift, using the evolution of the fraction of young and old SNe~Ia as
    predicted by \cite{rigault2020}, and assuming constant underlying stretch
    distribution for each age population \textbf{consisting} of Gaussian
    mixtures. We test our prediction against published samples \textbf{cut} to
    have marginal magnitude selection effects, so that any observed change is
    indeed of astrophysical and not observational in origin. \textbf{In this first
    study, there are indications} that the underlying SN~Ia stretch distribution
    is evolving as a function of redshift, and that the young/old drifting model
    is a better description of the data than any time-constant model, including
    the sample-based asymmetric distributions often used to correct Malmquist
    bias. The favored underlying stretch model is the bimodal one derived from
    \cite{rigault2020}, \textbf{composed of} a high-stretch mode shared by both
    young and old environments, and a low-stretch mode exclusive to old
    environments. The precise impact of the redshift evolution of the SN Ia
    population intrinsic properties on cosmology remains to be studied. Yet, the
    astrophysical drift of the SN stretch distribution does affect current
    Malmquist bias corrections and thereby distances derived from SN affected by
observational selection effects. We highlight that such a bias will increase with surveys
covering increasingly larger redshift ranges, which is particularly important
for LSST.}

@@

%For the last decades, numerous analyses have studied 
The relation between SNe~Ia
and their host galaxy properties has been studied extensively. The first key finding
was that the standardized SNe~Ia magnitudes
significantly depend on the host galaxy stellar mass, with SNe~Ia from high-mass host 
galaxies being
brighter on average \cite[e.g.][]{kelly2010, sullivan2010, childress2013,
betoule2014, rigault2020, kim19}. This mass-step correction is currently used in
cosmological analyses \citep[e.g.][]{betoule2014, scolnic2018a}, including for
deriving $H_0$ \citep{riess2016, riess2019}. Yet, the underlying connection
between the SNe and their host galaxies remains unclear, particularly when using global properties such
as the host stellar mass. This in turn raises the question of the accuracy of such
corrections since host galaxy global properties evolve with redshift. 
More recently, studies have used the local SN host galaxy environment to probe
more direct connections between the SNe and their host galaxy environment \citep{rigault2013},
showing that local age tracers such as the Local specific Star Formation Rate
(LsSFR) or the local color are more strongly correlated with the standardized SN
magnitude \citep{roman2018, kim18,rigault2020}. Further reinforcing this
connection is the identification of SN~Ia spectral features that are 
correlated with LsSFR \citep{nordin2018}. These results suggest age as the driving
parameter underlying the mass-step. If true, this would have a significant
impact for cosmology, since the environmental correction to apply to SN
standardization could strongly vary with redshift. \citep{rigault2013,
childress2014, scolnic2018a}. Yet, the importance of local SN environmental
studies remains highly debated \cite[e.g.][]{jones2015, jones2019} and
especially the impact of such an astrophysical bias has on the derivation of
$H_0$ \citep{jones2015, riess2016, riess2018, rose2019}.

@@

In this paper, we take a step aside \mr{from the cosmological analyses} to probe the validity of our modeling of the
SN population, which we claim to be constituted of two age-populations
\citep{rigault2013,rigault2015,rigault2020}: one old and one younger, the former
having on average lower lightcurve stretches and being brighter after
standardization. We use the correlation between the SN age, as probed by the
LsSFR, and the SN stretch to model the expected evolution of the underlying SN
stretch distribution as a function of redshift. This modeling relies on three
assumptions: (1) there are two distinct populations of SNe~Ia; (2) the relative
fraction of each of these populations as a function of redshift follows the
model presented in \cite{rigault2020} and (3) the underlying distribution of
stretch for each age sample is constant. This paper aims at testing this
specific model with datasets from the literature. 
% Note that the progenitor age
% as traced by the LsSFR seems to capture physical feature intrinsic to the
%progenitor and/or explosion mechanism that the stretch alone is not capturing
% \citep{nordin2018}.

@@

Assuming sufficient (and unbiased) spectroscopic follow-up for acquiring SN types
and host galaxy redshifts, the observation selection effects of magnitude-limited surveys should be
negligible below a given redshift at which even the faintest normal SNe~Ia can be
observed. \textbf{Aiding in the construction of nearly volume-limited 
subsamples is the fact that the SN~Ia population trails off towards fainter SNe~Ia.
A complication is that complete spectroscopic follow-up has not always been
the norm, as discussed below.}
In contrast, targeted surveys have highly complex observational
selection functions and so are discarded from our analysis. High-redshift SN cosmology
samples, such as Pantheon, are predominately from magnitude-limited surveys from
which volume-limited SN~Ia subsamples can be constructed.
%\textbf{which have easier to model observational selection functions}.

We present in Fig.~\ref{fig:maglim} the lightcurve stretch and color of SNe~Ia
from the following surveys: PanStarrs \citep[PS1][]{rest2014}, the Sloan Digital
Sky Survey \citep[SDSS][]{frieman2008} and the SuperNovae Legacy Survey
\citep[SNLS][]{astier2006}. An ellipse in the \textsc{\texttt{SALT2.4}} $(x_1,
c)$ plane with $x_1 = \pm 3$ and $c = \pm 0.3$ encapsulates the full
parent distribution \citep{guy2007,betoule2014}; see also \citet{bazin2011} and
\citet{campbell2013} for similar contours, the second using a more conservative
$|c| \leq 0.2$ cut. Assuming the SN absolute magnitude with $x_1=0$ and $c=0$ is
$M_0=-19.36$ \citep{kessler2009,scolnic2014}, we can derive the absolute
standardized magnitude at maximum of light $M = M_0 - \alpha x_1 + \beta c$
along the aforementioned ellipse given the standardization coefficient
$\alpha=0.156$ and $\beta=3.14$ from \cite{scolnic2018a}: the faintest SN~Ia is
that with $(x_1=-1.65, c=0.25)$ and an absolute standardized magnitude at peak
in Bessel $B$ band of $M^{t_0}_{\min} = -18.31$~mag. Since one ought to detect
this object typically 5 days before and a week after peak to build a suitable
lightcurve, the effective limiting standardized absolute magnitude is
approximately $M_{\lim} = -18.00$~mag. Hence, given the magnitude limit
$m_{\lim}$ of a magnitude limited survey, one can derive the maximum redshift
$z_{\lim}$ above which the faintest SNe~Ia will be missed using the relation
between apparent magnitude, redshift and absolute magnitude $\mu(z_{\lim}) =
m_{\lim} - M_{\lim}$.

@@

\textbf{We will thus consider a set of cuts that will define a first fiducial
    sample, taking the limits as initially suggested by the previous
    completeness analysis. \textbf{However,} as this solution might be an overly
    simplified way to create a complete sample (e.g., because it ignores spectroscopic follow-up
    in efficiency for redshifts below $z_{lim}$), we also consider another set of cuts to define a so-called
    ``conservative'' sample. The latter is smaller and therefore less statistically
    constraining, but also even less prone to observational selection effects. If the
    redshift drift is still significant in the conservative sample, it would be
    even more meaningful in a carefully-tailored selection-free sample. These
    samples are adequate for the goal
    of this study, which is to develop a first implementation of a model for drift in
    SN Ia properties. If fruitful, the sample selection can later be refined, if needed, with a
    more detailed
    %drift modeling before refining it on a more complex, yet more accurate,
%follow-up analysis also accounting for precise 
model of the observational selection, e.g., using
the SNANA package \citep{SNANA}.}

SNLS typically acquired SNe~Ia in the redshift range $0.4<z<0.8$; at these
redshifts, the rest-frame Bessel $B$ band roughly corresponds to the SNLS $i$
filter, which has a $5\sigma$ depth of
24.8~mag\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT
final release website.}}. This converts to a $z_{\lim}=0.60$, in agreement with
\cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}. Fig.~14 of
\citet[][see their Section~5]{perrett2010} suggests however a lower limit of
$z_{\lim}=0.55$. \textbf{We will therefore use $z=0.60$ and $z=0.55$ as redshift
limits for the fiducial and conservative samples, respectively, for SNLS.}

Similarly, PS1 observed SNe~Ia in the range $0.2<z<0.4$, their $g$-band
$5\sigma$ depth is 23.1~mag \citep{rest2014}, which yields $z_{\lim}=0.31$, in
agreement with, e.g., Fig.~6 of \cite{scolnic2018a}. \textbf{If we were to be
    conservative, this figure would also suggest of a more stringent
$z_{\lim}=0.27$ cut; hence, we will use $0.31$ and $0.27$ for our fiducial and
conservative samples, respectively, for PS1.}

In a similar redshift range, SDSS has a limiting magnitude of 22.5
\citep{dilday2008, sako2008}, which would lead to $z_{\lim}=0.24$. However, the
SDSS surveys \textbf{had to contend with limited spectroscopic resources. As
discussed in} \citet[][Section~2]{kessler2009}, during the first year of SDSS,
SNe~Ia with $r<20.5$~mag were favored for spectroscopic follow-up, corresponding
to a redshift cut at $0.15$. For the rest of the SDSS survey, additional
spectroscopic resources were available, and \cite{kessler2009} and
\cite{dilday2008} show a reasonable completeness up to $z_{\lim}=0.2$. Based on this,
\textbf{we will use $z_{\lim}=0.20$ and $z_{\lim}=0.15$ for our
fiducial and conservative samples, respectively, for SDSS.}

The sample selection is summarized in Table~\ref{tab:sample}, and the redshift
distribution of these three surveys is shown in Fig.~\ref{fig:cuts}. As
expected, the selected redshift limits \textbf{are roughly located slightly
before the peak of these histograms. In Section \ref{ssec:verify} we validate that these
redshift limits are effective for constructing nearly volume-limited subsamples
from samples that were initially more closely magnitude-limited in their search
or spectroscopic follow-up.}

\begin{table}
    \centering
    \caption{Composition of the SNe~Ia dataset used in this analysis.
    Conservative cuts are indicated in parentheses. \textbf{The SNf limit is set by
\cite{rigault2020}, see text.}}
    \label{tab:sample}
    \begin{tabular}{l c c}
        \hline\hline
        Survey & $z_{\lim}$ & $N_{\mathrm{SN}}$ \\
        \hline
        SNf & 0.08 & 114 \\
        SDSS & 0.20 (0.15) & 167 (82)\\
        PS1 & 0.31 (0.27) & 160 (122)\\
        SNLS & 0.60 (0.55) & 102 (78)\\
        HST & -- & 26 \\
        \hline
        Total & -- & 569 (422) \\
        \hline
    \end{tabular}
\end{table}

@@

In addition, we use the SNe~Ia from the Nearby Supernova Factory
\citep[SNfactory,][]{aldering2002} published in \cite{rigault2020} and that have
been discovered from non-targeted searches (114 SNe~Ia, see their sections~3
and~4.2.2; SNe~Ia time series are published in \citealt{saunders2020}, see also
\citealt{aldering2020}). \textbf{For this dataset, spectroscopic screening was
    done for candidates with $r \lesssim 19.5$; redshift cuts were then applied
    when selecting which SN Ia to follow, resulting in a redshift range of $0.02
    < z < 0.09$, further reduced to $<0.08$ in \cite{rigault2020} for extracting
local host properties.} These 114 SNfactory SNe~Ia are thus in the
volume-limited part of the survey (Aldering et al., in prep.), and are therefore
assumed to be a random sampling of the underlying SN population. The SNfactory
sample is particularly useful for studying SN property drift, as it enables us
to have a large complete SN~Ia sample at $z<0.1$.

Finally, we include the HST sample from Pantheon \citep{strolger04}.
\textbf{These high-redshift SNe are of great interest as they provide the
    greatest leverage for testing evolution. While at these redshifts the
    supernovae typing is challenging, the target classification was robust
    enough to include them within the cosmological analysis \citep{scolnic2018a}
    and we do not impose further cuts. Section~\ref{sec:results} highlights
    that, while compatible with it, our results are not dependent on the
    inclusion of this dataset.}

We present the stretch distribution and redshift histogram of these five surveys
up to their respective $z_{\lim}$ in Fig.~\ref{fig:sample}. \textbf{We observe here
that the fraction of low-stretch SNe (typically $x_1 < -1$) appears to decrease as
a function of redshift; this is confirmed in Fig.~\ref{fig:modelall}, in which}
the evolution of the mean stretch is shown, \textbf{with} the data split in
redshift bins of regular sample size. We see that SNe~Ia at higher redshift have
on average larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$) than those at lower
redshift ($-0.17\pm 0.10$ at $z\sim0.05$), suggesting that the underlying
stretch distribution is \textbf{evolving with redshift}.

@@

\textbf{In section~\ref{ssec:cuts}, we have built volume-limited samples from a
    set of magnitude-limited ones, using simple redshift cuts. This simplified
    approach is statistically sub-optimal, but should be enough to test
    our key question: whether redshift evolution of stretch is compatible with the
    \cite{rigault2020} model. However, the possibility remains that a complex
    observational selection function related to spectroscopic follow-up efficiencies below our fiducial (or
    even conservative) redshift cuts could still affect our sample, making it not fully
volume-limited; this would, in turn, bias our conclusion on the astrophysical
drift of the SNe~Ia population. We now look at this possibility.}

\textbf{To test for the existence of potential leftover observational selection biases in our
    sample, we compare the stretch and color distributions of the SNe~Ia
    originating from different datasets having overlapping redshift ranges: these
    distributions should be similar if they reflect the underlying parent population. 
    Note that the redshift range has to be
    narrow enough so that any drift would be negligible.}

\textbf{The two samples that overlap the most in redshift are PS1 and SDSS in
    the redshift range $0.10 < z < 0.20$ (see Fig~\ref{fig:sample}). This overlapping subsample
    consists of the 146 SNe~Ia at the high-redshift end of SDSS and thus
    the most likely to be affected by residual observational selection effects (see the
    corresponding discussion in section~\ref{ssec:cuts}). Over that same
    redshift range, PS1 has 52 SNe~Ia, which are in the lowest redshift bins and
    thus unlikely to have any observational selection issue. To identify potential
    inconsistency between the PS1 and SDSS subsamples, Fig.~\ref{fig:distrib}
    (upper panels) compares the stretch and color distribution of both these
    surveys. The resulting Kolmogorov-Smirnov (KS) similarity test $p$-values
($p >10\%$) do not support any inconsistency, in agreement with the
visual impression from Fig.~\ref{fig:distrib}.}

\textbf{We perform a similar analysis for PS1 and SNLS over the redshift range
    $0.20 < z < 0.31$ (Fig.~\ref{fig:distrib}, lower panels), where the same
    conclusion can be drawn: there is no substantial sign of discrepancy in the
    stretch and color distributions between the low- and high-end of our fiducial SNLS
    and PS1 samples, respectively. Nonetheless, the small size of the SNLS
    dataset at $z < 0.31$ (12 SNe~Ia vs. 90 for PS1) limits the sensitivity of
    this test, and only a strong deviation would be noticeable. Extending the
redshift range to $0.20 < z < 0.40$ (though we have no PS1 data above 0.3)
allows to increase the SNLS subsample to 31, yet the stretch $p$-values remains
high (34\%) showing no sign of inconsistency.}

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/both-cut_SDSS_SNLS_PS1.pdf}
    \caption{\textbf{$x_1$ (left) and $c$ (right) distribution histograms of
            different surveys overlapping in redshift. \textit{Facing up}: SDSS
            and PS1 within the redshift range $0.10 < z < 0.20$; \textit{facing
            down}: PS1 and SNLS within the redshift range $0.20 < z < 0.31$.
            Error bars show the Poisson noise. Our stretch ``Base'' model is
            illustrated in orange at the mean redshift of the redshift ranges,
            $0.15$ and $0.25$, respectively. Kolmogorov-Smirnov test $p$-values
            are indicated on the top (resp. bottom) of each panel showing no
            sign that the SDSS and PS1 (resp. PS1 and SNLS) $x_1$ and $c$
            distributions are not drawn from the same underlying
    distributions}}
    \label{fig:distrib}
\end{figure}

\textbf{We finally highlight that the SNe~Ia color is more prone to observational selection
    effects than stretch as illustrated in Fig.~\ref{fig:maglim}; see also e.g.,
    Fig.~3 of \cite{kessler2017}. Hence, since the comparison of color
    distributions shows no significant hint of leftover observational selection effect, it
    further supports our claim that our simple redshift-based selection criteria
    are sufficient to build the complete SNe~Ia samples required to test the
    redshift evolution of the stretch distribution.}

@@

\begin{equation}
    \label{eq:delta}
    \delta(z) = \left( K^{-1} \times (1+z)^{-2.8} +1 \right)^{-1}
\end{equation}
with $K=0.87$. This model is comparable to the evolution \textbf{subsequently}
predicted by \cite{childress2014} based on SN rates in galaxies depending on
their quenching time as a function of their stellar mass.

@@

The stretch probability distribution function (pdf) of a given SN will be the
linear combination of the stretch distributions of these two population weighted
by its probability $y^i$ to be young (see Section~\ref{sec:basemodelapplied}).
But generally, the fraction of young SNe~Ia as a function of redshift is given
by $\delta(z)$ (see Eq.~\ref{eq:delta}) and therefore, our redshift drift model
of the underlying \textbf{stretch} distribution of SNe~Ia as a function of redshift $X_1(z)$ is
given by:
\begin{align}\label{eq:stretchz}
    X_1(z) = \delta(z)&\times \mathcal{N}(\mu_1,\sigma_1{}^2) + \nonumber \\
    (1-\delta(z))&\times \left[ a\times\mathcal{N}(\mu_1,\sigma_1{}^2) +
    (1-a)\times\mathcal{N}(\mu_2,\sigma_2{}^2) \right]
\end{align}
\textbf{This is our Base drifting model.}

>>>>

\abstract{The detailed nature of type Ia supernovae (SNe~Ia) remains uncertain,
    and as survey statistics increase, the question of astrophysical systematic
    uncertainties rises, notably that of SN~Ia population evolution. In this
    paper, we study the dependence with redshift of the SN~Ia \texttt{SALT2.4}
    lightcurve stretch, a purely intrinsic SN property, to probe its potential
    redshift drift. The SN stretch has been shown to strongly correlate with the
    SN environment, notably with stellar age tracers. We model the underlying
    stretch distribution as a function of redshift, using the evolution of the
    fraction of young and old SNe~Ia as predicted by \cite{rigault2020}, and
    assuming constant underlying stretch distribution for each age population
    consisting of Gaussian mixtures. We test our prediction against published
    samples cut to have marginal magnitude selection effects, so that any
    observed change is indeed of astrophysical and not observational in origin.
    In this first study, there are indications that the underlying SN~Ia stretch
    distribution is evolving as a function of redshift, and that the young/old
    drifting model is a better description of the data than any time-constant
    model, including the sample-based asymmetric distributions often used to
    correct Malmquist bias \textbf{at a significance of over 5 $\sigma$}. The
    favored underlying stretch model is the bimodal one derived from
    \cite{rigault2020}, composed of a high-stretch mode shared by both young and
    old environments, and a low-stretch mode exclusive to old environments. The
    precise impact of the redshift evolution of the SN Ia population intrinsic
    properties on cosmology remains to be studied. Yet, the astrophysical drift
    of the SN stretch distribution does affect current Malmquist bias
    corrections and thereby distances derived from SN affected by observational
selection effects. We highlight that such a bias will increase with surveys
covering increasingly larger redshift ranges, which is particularly important
for LSST.}

@@

The relation between SNe~Ia and their host galaxy properties has been
studied extensively. The first key finding was that the standardized SNe~Ia
magnitudes significantly depend on the host galaxy stellar mass, with
SNe~Ia from high-mass host galaxies being brighter on average
\cite[e.g.][]{kelly2010, sullivan2010, childress2013, betoule2014, rigault2020,
kim19}. This mass-step correction is currently used in cosmological analyses
\citep[e.g.][]{betoule2014, scolnic2018a}, including for deriving $H_0$
\citep{riess2016, riess2019}. Yet, the underlying connection between the SNe and
their host galaxies remains , particularlyunclear when using
global properties such as the host stellar . This in turnmass raises
the question of the accuracy of such corrections since host galaxy
global properties evolve with redshift. More recently, studies have used the
local SN host galaxy environment to probe more direct connections
between the SNe and their host galaxy environments \citep{rigault2013},
showing that local age tracers such as the Local specific Star Formation Rate
(LsSFR) or the local color are more strongly correlated with the standardized SN
magnitude \citep{roman2018, kim18,rigault2020}. Further reinforcing this
connection is the identification of SN~Ia spectral features that are correlated
with LsSFR \citep{nordin2018}. These results suggest age as the driving
parameter underlying the mass-step. If true, this would have a significant
impact for cosmology, since the environmental correction to apply to SN
standardization could strongly vary with redshift. \citep{rigault2013,
childress2014, scolnic2018a}. Yet, the importance of local SN environmental
studies remains highly debated \cite[e.g.][]{jones2015, jones2019} and
especially the impact of such an astrophysical bias has on the derivation of
$H_0$ \citep{jones2015, riess2016, riess2018, rose2019}.

@@

In this paper, we take a step aside from the cosmological analyses to
probe the validity of our modeling of the SN population, which we claim to be
constituted of two age-populations \citep{rigault2013,rigault2015,rigault2020}:
one old and one younger, the former having on average lower lightcurve stretches
and being brighter after standardization. We use the correlation between the SN
age, as probed by the LsSFR, and the SN stretch to model the expected evolution
of the underlying SN stretch distribution as a function of redshift. This
modeling relies on three assumptions: (1) there are two distinct populations of
SNe~Ia; (2) the relative fraction of each of these populations as a function of
redshift follows the model presented in \cite{rigault2020} and (3) the
underlying distribution of stretch for each age sample is constant. This paper
aims at testing this specific model with datasets from the literature.

@@

Assuming sufficient (and unbiased) spectroscopic follow-up for acquiring
SN types and host galaxy redshifts, the observation selection
effects of magnitude-limited surveys should be negligible below a given redshift
at which even the faintest normal SNe~Ia can be observed.
Aiding in the construction of nearly volume-limited subsamples is the
    fact that the SN~Ia population trails off towards fainter SNe~Ia. A
    complication is that complete spectroscopic follow-up has not always been
    the norm, as discussed below. In contrast, targeted surveys have highly
    complex observational selection functions and so are
    discarded from our analysis. High-redshift SN cosmology samples,
    such as Pantheon, are predominately from magnitude-limited surveys from
which volume-limited SN~Ia subsamples can be constructed.

We present in Fig.~\ref{fig:maglim} the lightcurve stretch and color of SNe~Ia
from the following surveys: PanStarrs \citep[PS1][]{rest2014}, the Sloan Digital
Sky Survey \citep[SDSS][]{frieman2008} and the SuperNovae Legacy Survey
\citep[SNLS][]{astier2006}. An ellipse in the \textsc{\texttt{SALT2.4}} $(x_1,
c)$ plane with $x_1 = \pm 3$ and $c = \pm 0.3$ encapsulates the full
parent distribution \citep{guy2007, betoule2014}; see also
\citet{bazin2011} and \citet{campbell2013} for similar contours, the second
using a more conservative $|c| \leq 0.2$ cut. Assuming the SN absolute magnitude
with $x_1=0$ and $c=0$ is $M_0=-19.36$ \citep{kessler2009,scolnic2014}, we can
derive the absolute standardized magnitude at maximum of light $M = M_0 - \alpha
x_1 + \beta c$ along the aforementioned ellipse given the standardization
coefficient $\alpha=0.156$ and $\beta=3.14$ from \cite{scolnic2018a}: the
faintest SN~Ia is that with $(x_1=-1.65, c=0.25)$ and an absolute standardized
magnitude at peak in Bessel $B$ band of $M^{t_0}_{\min} = -18.31$~mag. Since one
ought to detect this object typically 5 days before and a week after peak to
build a suitable lightcurve, the effective limiting standardized absolute
magnitude is approximately $M_{\lim} = -18.00$~mag. Hence, given the magnitude
limit $m_{\lim}$ of a magnitude limited survey, one can derive the maximum
redshift $z_{\lim}$ above which the faintest SNe~Ia will be missed using the
relation between apparent magnitude, redshift and absolute magnitude
$\mu(z_{\lim}) = m_{\lim} - M_{\lim}$.

@@

We will thus consider a set of cuts that will define a first fiducial
    sample, taking the limits as initially suggested by the previous
    completeness analysis. However, as this solution might be an overly
    simplified way to create a complete sample, e.g., because it ignores
    spectroscopic follow-up in efficiency for redshifts below $z_{lim}$, we
    also consider another set of cuts to define a so-called ``conservative''
    sample. The latter is smaller and therefore less statistically constraining,
    but also even less prone to observational selection effects. If the redshift
    drift is still significant in the conservative sample, it would be even more
    meaningful in a carefully-tailored selection-free sample. These samples are
    adequate for the goal of this study, which is to develop a first
    implementation of a model for drift in SN Ia properties. If fruitful, the
    sample selection can later be refined, if needed, with a more detailed model
of the observational selection, e.g., using the SNANA package \citep{SNANA}.

SNLS typically acquired SNe~Ia in the redshift range $0.4<z<0.8$; at these
redshifts, the rest-frame Bessel $B$ band roughly corresponds to the SNLS $i$
filter, which has a $5\sigma$ depth of
24.8~mag\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT
final release website.}}. This converts to a $z_{\lim}=0.60$, in agreement with
\cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}. Fig.~14 of
\citet[][see their Section~5]{perrett2010} suggests however a lower limit of
$z_{\lim}=0.55$. We will therefore use $z=0.60$ and $z=0.55$ as redshift
limits for the fiducial and conservative samples, respectively, for SNLS.

Similarly, PS1 observed SNe~Ia in the range $0.2<z<0.4$, their $g$-band
$5\sigma$ depth is 23.1~mag \citep{rest2014}, which yields $z_{\lim}=0.31$, in
agreement with, e.g., Fig.~6 of \cite{scolnic2018a}. If we were to be
    conservative, this figure would also suggest of a more stringent
$z_{\lim}=0.27$ cut; hence, we will use $0.31$ and $0.27$ for our fiducial and
conservative samples, respectively, for PS1.

In a similar redshift range, SDSS has a limiting magnitude of 22.5
\citep{dilday2008, sako2008}, which would lead to $z_{\lim}=0.24$. However, the
SDSS surveys had to contend with limited spectroscopic resources. As
discussed in \citet[][Section~2]{kessler2009}, during the first year of SDSS,
SNe~Ia with $r<20.5$~mag were favored for spectroscopic follow-up, corresponding
to a redshift cut at $0.15$. For the rest of the SDSS survey, additional
spectroscopic resources were available, and \cite{kessler2009} and
\cite{dilday2008} show a reasonable completeness up to $z_{\lim}=0.2$.
Based on this, we will use $z_{\lim}=0.20$ and $z_{\lim}=0.15$ for our
fiducial and conservative samples, respectively, for SDSS.

The sample selection is summarized in Table~\ref{tab:sample}, and the redshift
distribution of these three surveys is shown in Fig.~\ref{fig:cuts}. As
expected, the selected redshift limits are roughly located slightly
    before the peak of these histograms. In Section \ref{ssec:verify} we
validate that these redshift limits are effective for constructing nearly
volume-limited subsamples from samples that were initially more closely
magnitude-limited in their search or spectroscopic follow-up.

\begin{table}
    \centering
    \caption{Composition of the SNe~Ia dataset used in this analysis.
    Conservative cuts are indicated in parentheses. The SNf limit is set
by \cite{rigault2020}, see text.}
    \label{tab:sample}
    \begin{tabular}{l c c}
        \hline\hline
        Survey & $z_{\lim}$ & $N_{\mathrm{SN}}$ \\
        \hline
        SNf & 0.08 & 114 \\
        SDSS & 0.20 (0.15) & 167 (82)\\
        PS1 & 0.31 (0.27) & 160 (122)\\
        SNLS & 0.60 (0.55) & 102 (78)\\
        HST & -- & 26 \\
        \hline
        Total & -- & 569 (422) \\
        \hline
    \end{tabular}
\end{table}

@@

In addition, we use the SNe~Ia from the Nearby Supernova Factory
\citep[SNfactory,][]{aldering2002} published in \cite{rigault2020} and that have
been discovered from non-targeted searches (114 SNe~Ia, see their sections~3
and~4.2.2; SNe~Ia time series are published in \citealt{saunders2020}, see also
\citealt{aldering2020}). For this dataset, spectroscopic screening was
    done for candidates with $r \lesssim 19.5$; redshift cuts were then applied
    when selecting which SN Ia to follow, resulting in a redshift range of $0.02
< z < 0.09$, further reduced to $<0.08$ in \cite{rigault2020} for extracting
local host properties. These 114 SNfactory SNe~Ia are thus in the
volume-limited part of the survey (Aldering et al., in prep.), and are therefore
assumed to be a random sampling of the underlying SN population. The SNfactory
sample is particularly useful for studying SN property drift, as it enables us
to have a large complete SN~Ia sample at $z<0.1$.

Finally, we include the HST sample from Pantheon \citep{strolger04}.
These high-redshift SNe are of great interest as they provide the
    greatest leverage for testing evolution. While at these redshifts the
    supernovae typing is challenging, the target classification was robust
    enough to include them within the cosmological analysis \citep{scolnic2018a}
    and we do not impose further cuts. Section~\ref{sec:results} highlights
    that, while compatible with it, our results are not dependent on the
    inclusion of this dataset.

We present the stretch distribution and redshift histogram of these five surveys
up to their respective $z_{\lim}$ in Fig.~\ref{fig:sample}. We observe here
that the fraction of low-stretch SNe (typically $x_1 < -1$) appears to decrease as
a function of redshift; this is confirmed in Fig.~\ref{fig:modelall}, in which
the evolution of the mean stretch is shown, with the data split in
redshift bins of regular sample size. We see that SNe~Ia at higher redshift have
on average larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$) than those at lower
redshift ($-0.17\pm 0.10$ at $z\sim0.05$), suggesting that the underlying
stretch distribution is evolving with redshift.

@@

In section~\ref{ssec:cuts}, we have built volume-limited samples from a set of
magnitude-limited ones, using simple redshift cuts. This simplified approach is
statistically sub-optimal, but should be enough to test our key question:
whether redshift evolution of stretch is compatible with the \cite{rigault2020}
model. However, the possibility remains that a complex observational selection
function related to spectroscopic follow-up efficiencies below our fiducial (or
even conservative) redshift cuts could still affect our sample, making it not
fully volume-limited; this would, in turn, bias our conclusion on the
astrophysical drift of the SNe~Ia population. We now look at this possibility.

To test for the existence of potential leftover observational selection biases
in our sample, we compare the stretch and color distributions of the SNe~Ia
originating from different datasets having overlapping redshift ranges: these
distributions should be similar if they reflect the underlying parent
population. We note that the redshift range has to be narrow enough so that any
drift would be negligible.

The two samples that overlap the most in redshift are PS1 and SDSS in the
redshift range $0.10 < z < 0.20$ (see Fig~\ref{fig:sample}). This overlapping
subsample consists of the 146 SNe~Ia at the high-redshift end of SDSS and thus
the most likely to be affected by residual observational selection effects (see
the corresponding discussion in section~\ref{ssec:cuts}). Over that same
redshift range, PS1 has 52 SNe~Ia, which are in the lowest redshift bins and
thus unlikely to have any observational selection issue. To identify potential
inconsistency between the PS1 and SDSS subsamples, Fig.~\ref{fig:distrib} (upper
panels) compares the stretch and color distribution of both these surveys. The
resulting Kolmogorov-Smirnov (KS) similarity test $p$-values ($p >10\%$) do not
support any inconsistency, in agreement with the visual impression from
Fig.~\ref{fig:distrib}.

We perform a similar analysis for PS1 and SNLS over the redshift range $0.20 < z
< 0.31$ (Fig.~\ref{fig:distrib}, lower panels), where the same conclusion can be
drawn: there is no substantial sign of discrepancy in the stretch and color
distributions between the low- and high-end of our fiducial SNLS and PS1
samples, respectively. Nonetheless, the small size of the SNLS dataset at $z <
0.31$ (12 SNe~Ia vs. 90 for PS1) limits the sensitivity of this test, and only a
strong deviation would be noticeable. Extending the redshift range to $0.20 < z
< 0.40$ (though we have no PS1 data above 0.3) allows to increase the SNLS
subsample to 31, yet the stretch $p$-values remains high (34\%) showing no sign
of inconsistency.

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/both-cut_SDSS_SNLS_PS1.pdf}
    \caption{$x_1$ (left) and $c$ (right) distribution histograms of
            different surveys overlapping in redshift. \textit{Facing up}: SDSS
            and PS1 within the redshift range $0.10 < z < 0.20$; \textit{facing
            down}: PS1 and SNLS within the redshift range $0.20 < z < 0.31$.
            Error bars show the Poisson noise. Our stretch ``Base'' model is
            illustrated in orange at the mean redshift of the redshift ranges,
            $0.15$ and $0.25$, respectively. Kolmogorov-Smirnov test $p$-values
            are indicated on the top (resp. bottom) of each panel showing no
            sign that the SDSS and PS1 (resp. PS1 and SNLS) $x_1$ and $c$
            distributions are not drawn from the same underlying
    distributions}
    \label{fig:distrib}
\end{figure}

We finally highlight that the SNe~Ia color is more prone to
    observational selection effects than stretch as illustrated in
    Fig.~\ref{fig:maglim}; see also e.g., Fig.~3 of \cite{kessler2017}. Hence,
    since the comparison of color distributions shows no significant hint of
    leftover observational selection effect, it further supports our claim that
    our simple redshift-based selection criteria are sufficient to build the
    complete SNe~Ia samples required to test the redshift evolution of the
stretch distribution.

@@

\begin{equation}
    \label{eq:delta}
    \delta(z) = \left( K^{-1} \times (1+z)^{-2.8} +1 \right)^{-1}
\end{equation}
with $K=0.87$. This model is comparable to the evolution subsequently
predicted by \cite{childress2014} based on SN rates in galaxies depending on
their quenching time as a function of their stellar mass.

@@

The stretch probability distribution function (pdf) of a given SN will be the
linear combination of the stretch distributions of these two population weighted
by its probability $y^i$ to be young (see Section~\ref{sec:basemodelapplied}).
But generally, the fraction of young SNe~Ia as a function of redshift is given
by $\delta(z)$ (see Eq.~\ref{eq:delta}) and therefore, our redshift drift model
of the underlying stretch distribution of SNe~Ia as a function of redshift $X_1(z)$ is
given by:
\begin{align}\label{eq:stretchz}
    X_1(z) = \delta(z)&\times \mathcal{N}(\mu_1,\sigma_1{}^2) + \nonumber \\
    (1-\delta(z))&\times \left[ a\times\mathcal{N}(\mu_1,\sigma_1{}^2) +
    (1-a)\times\mathcal{N}(\mu_2,\sigma_2{}^2) \right]
\end{align}
This is our Base drifting model.
